---
title: "AE 03: Duke Forest + data visualization"
author: "Derek Dubrule"
date: "Feb 1st, 2026"
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
```

## Exercise 1

Suppose you're helping some family friends who are looking to buy a house in Duke Forest.
As they browse Zillow listings, they realize some houses have garages and others don't, and they wonder: **Does having a garage make a difference?**

Luckily, you can help them answer this question with data visualization!

-   Make histograms of the prices of houses in Duke Forest based on whether they have a garage.
    -   In order to do this, you will first need to create a new variable called `garage` (with levels `"Garage"` and `"No garage"`).
    -   Below is the code for creating this new variable. Here, we `mutate()` the `duke_forest` data frame to add a new variable called `garage` which takes the value `"Garage"` if the text string `"Garage"` is detected in the `parking` variable and takes the test string `"No garage"` if not.

```{r}
#| label: garage
#| eval: false

duke_forest |>
  mutate(garage = if_else(str_detect(parking, "Garage"),   "Garage", "No garage"))
```

-   Then, facet by `garage` and use different colors for the two facets.
-   Choose an appropriate binwidth and decide whether a legend is needed, and turn it off if not.
-   Include informative title and axis labels.
-   Finally, include a brief (2-3 sentence) narrative comparing the distributions of prices of Duke Forest houses that do and don't have garages. Your narrative should touch on whether having a garage "makes a difference" in terms of the price of the house.

```{r}
#| label: price-garage

# add code here

duke_forest |>
  mutate(
    garage = if_else(str_detect(parking, "Garage"),   "Garage", "No garage")
    ) |>
  ggplot(aes(x = price, fill = garage)) +
  geom_histogram(binwidth = 40000) +
  facet_wrap(~ garage) +
  labs(
    title = "House Prices in Duke Forest based on having a garage",
    x = "House Price",
    y = "Number of houses"
  ) +
  theme(legend.position = "none")

```

Add narrative here...

Houses that include a garage appear to be listed at a higher price more often then houses with no garage.
But that being said there is a lot of overlap between the two graphs - so having a garage or not is not the 
only driver in house price. So having a garage makes a difference to a certain extent - but it is not the 
only factor when considering house prices.

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::

## Exercise 2

It's expected that within any given marker larger houses will be priced higher.
It's also expected that the age of the house will have an effect on the price.
However in some markets new houses might be more expensive while in others new construction might mean "no character" and hence be less expensive.
So your family friends ask: "In Duke Forest, do houses that are bigger and more expensive tend to be newer ones than those that are smaller and cheaper?"

Once again, data visualization skills to the rescue!

-   Create a scatter plot to exploring the relationship between `price` and `area`, conditioning for `year_built`.
-   Use `geom_smooth()` with the argument `se = FALSE` to add a smooth curve fit to the data and color the points by `year_built`.
-   Include informative title, axis, and legend labels.
-   Discuss each of the following claims (1-2 sentences per claim). Your discussion should touch on specific things you observe in your plot as evidence for or against the claims.
    -   Claim 1: Larger houses are priced higher.
    -   Claim 2: Newer houses are priced higher.
    -   Claim 3: Bigger and more expensive houses tend to be newer ones than smaller and cheaper ones.

```{r}
#| label: price-area

# add code here
ggplot(duke_forest, aes(x = area, y = price, color = year_built)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  scale_color_gradient(low = "green", high = "red") +
  labs(
    title = "House price in comparison to the area in Duke Forest",
    x = "Area in sq ft",
    y = "Price $",
    color = "Year Built"
  )
```

Add narrative here...

claim 1:
The plot shows a positive relationship between area and price: as area increases, price usually increases as well.
This is highlighted through the smooth curve as there is a clear trend upwards.

claim 2:
Newer homes with a higher year built age(orange / red) appear more often among the higher priced listings.
But this pattern is not a clear positive relationship, there are still older houses that are expensive and new
houses that cost less.

claim 3:
The majority of the biggest and most expensive homes are colored as a newer in comparison to the
smaller and less expensive homes which tend to be older. So this supports the claim that Bigger and 
more expensive houses tend to be newer ones than smaller and cheaper ones.


::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::
